---
title: "Silent Spring Text Analysis"
description: |
  A text and sentiment analysis of Rachel Carson's Silent Spring.
author:
  - name: Aaron Kreisberg
    url: {}
date: 2021-03-15
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


# Text Analysis-Silent Spring by Rachel Carson

### Introduction 

Silent Spring by Rachel Carson is one the seminal texts of 20th century. It brought greater environmental awareness regarding chemicals and their impacts. A text and sentiment analysis of Silent Spring can yield some insight on this important book. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # Read in the tidyverse
library(here) # Read in the here package
library(tidytext) # Read in tidytext package
library(textdata) # Read in textdata package
library(pdftools) # Read in pdftools package
library(ggwordcloud) # Read in ggwordcloud
library(jpeg) # Read in jpeg
library(grid) # Read in grid

```

Silent Spring is available for download as free PDF online. Below is page 47 of Silent Spring after reading in the downloaded .pdf using pdf_text(here()).

```{r}

silent_spring_text <- pdf_text(here("_posts", "2021-03-12-silentspringtextanalysis", "data", "silent_spring.pdf"))


silent_spring_p47 <- silent_spring_text[47]
silent_spring_p47

```

Before pursing further textual analysis, the text of Silent Spring needs to be converted into a "data.frame()" and the fact that lines are separated by \r also needs to be addressed. Using "str_split()" the text can be broken into individual lines, and further wrangled for analysis using "unnest()" and "mutate(text_full = str_trim())".

```{r}
silent_spring_tidy <- data.frame(silent_spring_text) %>% 
  mutate(text_full = str_split(silent_spring_text, pattern = '\\r')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
```

So this textual anlysis is concentrating on the body of Silent Spring, the introduction and post-body text can be removed using "slice()".

```{r}
silent_spring_df <- silent_spring_tidy %>% 
  slice(201:6201) 
     
  
```

The text of Silent Spring is converted into tokens with "unnest_tokens()". The number of times words appear is the determined using "count()" and stop words (but, and, the, etc.) are removed using "anti_join(stop_words)".

```{r}
silent_spring_tokens <- silent_spring_df %>% 
  unnest_tokens(word, text_full) %>% 
  dplyr::select(-silent_spring_text)

silent_spring_wordcount <- silent_spring_tokens %>% 
  count(word)

silent_spring_nonstop_words <- silent_spring_tokens %>% 
  anti_join(stop_words)


nonstop_counts <- silent_spring_nonstop_words %>% 
  count(word)
```

To get the top 100 words "arrange() %>% slice(1:100)". The top two words are "chemicals" and "chemical", with "ddt" and "spraying" also making the top 5. 

```{r}
top100 <- nonstop_counts %>% 
  arrange(-n) %>% 
  slice(1:100)
```

```{r}
ddt_background <- readJPEG(here("_posts", "2021-03-12-silentspringtextanalysis", "data", "ddt_spraying.jpg"))
```


### Visualizations 


A word cloud of the top 100 words is produced using "ggplot() + geom_text_wordcloud".

```{r}
silent_spring_cloud <- ggplot(data = top100, aes(label = word)) +
  annotation_custom(rasterGrob(ddt_background, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                    -Inf, Inf, -Inf, Inf) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "pentagon") +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c("red", "yellow")) +
  theme_minimal()

silent_spring_cloud
```

**Figure 1**: Silent Spring Wordcloud. Data: **Silent Spring**, Rachel Carson (1962)

```{r}
top20 <- top100 %>% 
  slice(1:20) %>% 
  group_by(-n) 

ddt_for_me <- readJPEG(here("_posts", "2021-03-12-silentspringtextanalysis", "data", "ddt1.jpg"))
```


A column graph of the Top 20 words in Silent Spring.

```{r}
silent_spring_column <- ggplot(data = top20, aes(x = reorder(word, n), y = n)) +
  annotation_custom(rasterGrob(ddt_for_me, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                    -Inf, Inf, -Inf, Inf) +
  geom_col(aes(alpha = 0), show.legend = FALSE) +
  scale_colour_manual(
    values = cols,
    aesthetics = c("colour", "fill")
  ) +
  coord_flip() +
  labs(title = "Top 20 Word in Silent Spring by Rachel Carson",
       y = "Count",
       x = "Word") +
  theme_classic()
  
  

silent_spring_column
```

**Figure 2**: Silent Spring Top 20 Words Column Plot. Data: **Silent Spring**, Rachel Carson (1962)


NRC sentiment analysis with "inner_join(get_sentiments())".

```{r, message = FALSE}
silent_spring_nrc <- silent_spring_nonstop_words %>% 
  inner_join(get_sentiments("nrc"))

silent_spring_nrc_counts <- silent_spring_nrc %>% 
  count(sentiment)

ddt_plane <- readJPEG(here("_posts", "2021-03-12-silentspringtextanalysis", "data", "ddt_plane.jpg"))

ggplot(data = silent_spring_nrc_counts, aes(x = sentiment, y = n, color = "red")) +
  annotation_custom(rasterGrob(ddt_plane, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                    -Inf, Inf, -Inf, Inf) +
  geom_col(aes(alpha = 0.5), show.legend = FALSE) +
  labs(title = "Silent Spring NRC Sentiment Analysis",
       x = "Sentiment",
       y = "Word Count")
```



**Figure 3**: Silent Spring NRC Sentiment Analysis. Data: **Silent Spring**, Rachel Carson (1962)


### Citation

Data: **Silent Spring**, Rachel Carson (1962)



