---
title: "a3_task3_aaron_kreisberg"
author: "Aaron Kreisberg"
date: "11/06/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

First the urchins.csv file will be read in using read.csv(here("data", "urchins.csv")) from the data folder in the R.project, and names will be cleaned using clean_names()

```{r}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Section 1 - Size Comparison of purple and red urchins

# Exploratory data visualization of purple and red urchins

First the distribution of sizes of purple and red urchins will be examined visually. This will be done by using gg_plot, with both a faceted histogram (ggplot() + geom_histogram + facet_wrap(~common_name)) and a boxplot (ggplot() + geom_boxplot())

```{r}
ggplot(data = urchins, aes(x = size, fill = common_name)) +
  geom_histogram(color = "black") +
  facet_wrap(~common_name) +
  scale_fill_manual(values = c("purple", "red")) +
  theme(legend.position = "none") +
  labs(x = "Urchin Size (diameter to the nearest 0.5 cm)", y = "# of Urchins")
```
```{r}
ggplot(data = urchins, aes(x = common_name, y = size, fill = common_name)) +
  geom_boxplot() +
  scale_fill_manual(values = c("purple", "red")) +
  theme(legend.position = "none") +
  labs(x = "Urchins", y ="Urchin Size (nearest 0.5 cm)")
 
  
```

# Descriptive statistics table of purple and red urchins data

Next the data will be explored in a summary statistics table by using goup_by(common_name) followed by summarise(mean(), median (), sd(), and n()). The kbl() from the kableExtra package will improve the aesthetics of the summary statistics table

```{r}
urchin_stats <- urchins %>%
  group_by(common_name) %>% 
  summarise(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size), 
            n = n()) 
kbl(urchin_stats) %>% 
  kable_classic(full_width = TRUE)

actual_difference = (6.776489-4.469797)
actual_difference
percent_difference = ((6.776489-4.469797)/((6.776489+4.469797)/2))*100
percent_difference
```
 

Based on the above exploratory graphs and table, both populations appear to relatively normally distributed with purple maybe showing a slight positive skew and red urchins showing more upper range outliers. Based on summary statistics, red urchins show a greater mean and median size, and a larger standard deviation, although there is a greater number of purple urchins (n = 31805) versus red urchins (n = 27408). The actual difference between mean urchin diameter is ~2.31, while the percent difference (calculated in the code section above) is ~41%; the mean red urchin is 41% larger in diameter than the mean purple urchin which can be a more useful comparison then just a actual difference depending on the comparison.   

## Section 2 - Exploration of purple urchin counts by site and year

In a piped sequence of code a subset of the urchins data will be created. A new column is created to convert the date to class "Date" and creating a colunm with just the year.

```{r}
purple_urchins <- urchins %>%
    mutate(date_new = mdy(date)) %>% # Convert the date to class "Date" ~mdy(date)
  filter(common_name == "Purple Urchin") %>% # Filter for only the purple urchins ~filter()
  mutate(year = year(date_new)) %>% # Create a column with just the date ~mutate(year())
  mutate(site_full = case_when( # Create a new column with full names for sites case_when() 
    site == "NAPL" ~ "Naples Reef", 
    site == "AQUE" ~ "Arroyo Quemado Reef", 
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))

# Create a summary table of # of purple urchins by site and year, using the group_by() and count()
purple_urchin_site_year <- purple_urchins %>% 
  group_by(year, site_full) %>% 
  count(site_full)
  
purple_urchin_site_year  
```

```{r}
# Create an exploratory graph of counts of purple urchins at each site by year. This was done with ggplot() + geom_col() + facet_wrap()

ggplot(data = purple_urchin_site_year, aes(x = year, y = n, fill = site_full)) +
  geom_col() +
  facet_wrap(~site_full)


```

