---
title: "Spatial Analysis"
description: |
  Spatial analysis of 2008 California oil spills and California cetacean species richness.
author:
  - name: Aaron Kreisberg
date: 2021-02-24
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
    
---

# Part One-Spatial Exploration of 2008 California Oil Spills

### Introduction

In this document 2008 oil spills across California will be visualized using R. Information is from California Department of Fish and Wildlife (CDFW). 

```{r setup, echo = TRUE, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #Read in the tidyverse
library(here) #Read in the here package
library(janitor) #Read in the janitor package
library(sf) #Read in sf
library(tmap) #Read in tmap
library(raster) # Read in raster  **NOTE: Attach this BEFORE tidyverse**
library(fasterize) # Read in fasterize
library(rnaturalearth) # Read in rnaturalearth
```


Data from CDFW is read in and stored as "ca_spills". This data can be visualized in conjunction with a data that represents California's counties and their borders (CA_Counties_TIGER2016). "CA_Counties_TIGER2016" is stored as "ca_counties".

```{r, message = FALSE, warning = FALSE}
ca_spills <- read_sf(here("_posts", "2021-02-24-welcome", "data", "ds394"), layer = "ds394") %>% 
  clean_names()

ca_counties <- read_sf(here("_posts", "2021-02-24-welcome", "data", "ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  dplyr::select(name)

ca_counties <- st_transform(ca_counties, st_crs(ca_spills)) #Convert ca_counties from NAD84 to NAD83 so CRS matches
```

It is important when when working with spatial data that the Coordinate Reference System (CRS) of datasets match, otherwise it can hinder visualization and results. To accomplish this, st_transform() is used to convert the CRS of "ca_counties" from NAD84 to NAD83. The CRS can be double checked with st_crs().

```{r, include = FALSE}
st_crs(ca_spills) #NAD83

st_crs(ca_counties) #NAD83
```

### Exploratory Data Visualization

Using ggplot() and geom_sf() oil spills in California can be visualized.

```{r}

# Visualize with ggplot() and geom_sf():
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = ca_spills) +
  labs(title = "2008 California Oil Spills")

```
**Figure 1**: 2008 Oil Spills in California. Data: California Department of Fish and Wildlife 


### Interactive Visualization

Using the tmap package, an interactive visualization can be produced. Numbers for data points represent DFG Control numbers (CAFW was previously California Deparmtent of Fish and Game). OES is the Office of Emergency Services and each spill has OES number.

```{r, message = FALSE}
tmap_mode("view")

tm_shape(ca_spills) +
  tm_dots(col = "inlandmari",
          popup.vars = c("Spill Date:" = "dateofinci",
                         "OES Number:" = "oesnumber",
                         "City:" = "localecity",
                         "County:" = "localecoun"),
          title = "Spill Type")
```
**Figure 2**: Interactive Map of 2008 California Oil Spills. Spills are classified as either marine or inland. Spills are further classified by Office of Emergency Services number. City and county of occurrence are also noted. Data: California Department of Fish and Wildlife 



### Chloropleth Map of Inland Spills by County

To examine spills classified as inland spills (not occurring on the ocean or  waterbodies connected to the ocean), the "ca_spills" can be filtered using the filter(). This produces a dataframe called "ca_inlandspills" To create a chloropleth map, the "ca_inlandspills" dataframe is joined with the "ca_counties" dataframe using st_join(). Spills and counted by county using the count().

```{r}
 ca_inlandspills <- ca_spills %>% 
  filter(inlandmari == "Inland") 

ca_inlandspills_join <- ca_counties %>% 
  st_join(ca_inlandspills)

ca_inlandspills_counts <- ca_inlandspills_join %>% 
  count(name)


```


```{r}
ggplot() +
  geom_sf(data = ca_inlandspills_counts, aes(fill = -n, group = name)) +
  theme_minimal() +
  labs(
    title = "2008 Inland Oil Spills in California by County",
    fill = "Number of 2008 oil spills",
    caption = "Data: CA Department of Fish and Wildlife")
```

**Figure 3**: Chloropleth Map of 2008 Inland California Oil Spills by County.  Data: California Department of Fish and Wildlife

### Conclusions

Some conclusions from the above visualizations:

-2008 oil spills in California were concentrated in areas with major ports, particularly the Bay Area and Los Angeles county.

-There was a higher quantity of inland oil spills compared to oil spills classified as marine.

### Citation

Oil Spill Incident Tracking [ds394], Edition 2008, Publication date 2009-07-23. Marine data are entries from OSPR Oil Spill Prevention Specialists (OSPS) and is accurate to 98%. Inland data is based on California Emergency Management Agency (Cal EMA) reports and is only as accurate as the reporting party.Credits: Cal EMA, DFG-OSPR


# Part Two-Spatial Exploration of Coastal California Cetacean Species Richness

### Introduction

35 rasters that correspond to various cetacean species that are found off the California coast are used to explore species richness. Raster data layers are from AquaMaps (https://www.aquamaps.org/search.php).   


### Cetacean Rasters

The AquaMaps cetecean raster data layers are read in using "list.files()". Rather than reading in 35 individual raster layers, raster::stack() can be used to create a single data frame containing the rasters representing cetacean species. 

```{r, include = FALSE}
ca_cetacean_file <- list.files(here("_posts", "2021-02-24-welcome", "data", "ca_cetaceans"), full.names = TRUE)

cetacean_stack <- raster::stack(ca_cetacean_file)

cetacean_stack
```


### Threshold Model

The raster layers present the probability of the presence of a cetacean between zero and one within cells covering a geographic range of California's coast. This range covers the San Francisco Bay Area down to the California-Mexico border. By creating a threshold function that utilizes these probabilities, the rasters can be converted to display whether a cetacean is considered present or not at a particular area. For the purposes of this analysis a probability of 0.6 is utilized to determine whether a cetacean species is present or not. The "calc()" is used in conjunction with the created threshold function to create presence rasters for the cetacean species. 

```{r}
is_cetacean <- function(x, thresh = .6) {
  y <- ifelse(x >= thresh, 1, 0)
  return(y)
}

ca_cetacean_presence <- calc(cetacean_stack, fun = is_cetacean)

plot(ca_cetacean_presence, col = "blue")

```

**Figure 1**: Plots of AquaMaps Cetacean Rasters. Data: AquaMaps-Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016)


An additional usage of the "calc()" sums the presence rasters into a single raster that can be used for visualization. Before visualizing the raster it must be converted into a dataframe which is accomplished using "rasterToPoints() %>% as.data.frame()"  

```{r}


cetacean_richness <- calc(ca_cetacean_presence, fun = sum, na.rm = TRUE)

cetacean_df <-  raster::rasterToPoints(cetacean_richness) %>%
  as.data.frame()

plot(cetacean_richness)

```

**Figure 2**: Raster of Cetacean Species Richness. Data: AquaMaps-Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016)


### Cetacean Species Richness Visualization 

A visualization of cetacean species richness can be created using the dataframe of cetacean species richness. Before visualizing species richness an map of California's coast is needed. This can be read in from the rnaturalarth package using "ne_download()". The spatial dataframe is filtered to provide California. 


```{r, cache= TRUE, include = FALSE}
california <- ne_download(scale = "large", type = "states", category = "cultural", returnclass = "sf") %>% 
  filter(name %in% c("California", "Nevada"))
```

```{r}
ggplot(data = california) +
  geom_sf() 
```

**Figure 3** Map of California and Nevada. Data: Natural Earth (https://www.naturalearthdata.com/)


```{r}
ggplot() +
geom_raster(data = cetacean_df, aes(x = x, y = y, fill = layer)) +
  geom_sf(data = california) +
  coord_sf(xlim = c(-125,-116), ylim = c(32,38)) +
  labs(
    title = "California Cetacean Species Richness",
    fill = "Number of Cetacean Species",
    x = "Longitude",
    y = "Latitude") +
  theme_classic() +
  theme(panel.background = element_rect(fill = 'black'))
```

**Figure 4**: California Cetacean Species Richness. Species richness was determined by combining individual rasters into a single raster and converting this raster into a dataframe. Species richness is greatest near the coast and to the South. Species presence was determined using a function with a 0.6 threshold, with a 0.6 or greater chance of occurrence classified as the species being present and less then 0.6 classified as non-occurrence. Data: AquaMaps-Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016)




### Citation

Kaschner, K., K. Kesner-Reyes, C. Garilao, J. Segschneider, J. Rius-Barile, T. Rees, and R. Froese. 2019. AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 10/2019.
