---
title: "kreisberg_warblers"
author: "Aaron Kreisberg"
date: "12/10/2020"
output:
  html_document:
    code_folding: hide
---
# Exploration of Black-Throated Blue Warblers at Hubbard Brook Experimental Forest, 1986-2010

## Set-up and Data Wrangling

### Set-up Code Chunk

Various packages are read in to support exploration and visualization of the black-throated blue warbler data

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse) #tidyverse is read in with library()
library(here) #here is read in with library()
library(GGally) #GGally is read in with library()
library(stargazer) #stargazer is read in with library()
library(car) #car is read in with library()
library(broom) #broom is read in with library()
```

### Data Wrangling

Warbler data frame is wrangled using the filter() from the dplyr package  

```{r, include=FALSE}
warbler_data <- read_csv(here("exam_data", "bt_warblers.csv")) #Read in the warbler data using here()

warblers_winglength_mass <-warbler_data %>% 
  dplyr::filter(mass != "-99.99") %>% 
  dplyr::filter(wing_r != "-99.99") #The warbler data is filtered to remove missing values from the 'mass' and 'wing_r' data frames (filter(' ' != "-99.99"))

warblers_filtered <- warbler_data %>% 
  dplyr::filter(mass != "-99.99") %>% 
  dplyr::filter(wing_r != "-99.99") %>% 
  dplyr::filter(bill_depth != "-99.99") %>% 
  dplyr::filter(bill_culmen != "-99.99") %>% 
  dplyr::filter(tarsus1 != "-99.99")  #The warbler data is filtered to remove missing values from the all numeric data frames: 'bill_depth', 'bill-culmen', 'mass', 'tarsus', and 'wing_r' (filter(' ' != "-99.99"))
```

## Exploratory Data Visualizations

### Mass versus Wing Length

Warbler wing length is explored in relation to warbler mass, using various exploratory visualizations. A visualization of warbler length in relation to bird mass is displayed, with

```{r, include=FALSE}
#Exploratory histograms and qqplots to examine tendency, normality, and relationships between variables.
ggplot(data = warblers_winglength_mass, aes(x = mass)) +
  geom_histogram() 

ggplot(data = warblers_winglength_mass, aes(sample = mass)) +
  geom_qq() 
  

ggplot(data = warblers_winglength_mass, aes(x = mass, color = sex)) +
  geom_histogram() +
  facet_wrap(~age_banded)

ggplot(data = warblers_winglength_mass, aes(sample = mass, color = sex)) +
  geom_qq() +
  facet_wrap(~age_banded)

ggplot(data = warblers_winglength_mass, aes(x = wing_r)) +
  geom_histogram() 

ggplot(data = warblers_winglength_mass, aes(sample = wing_r)) +
  geom_qq() 
  

ggplot(data = warblers_winglength_mass, aes(x = wing_r, color = sex)) +
  geom_histogram() +
  facet_wrap(~age_banded)

ggplot(data = warblers_winglength_mass, aes(sample = wing_r, color = sex)) +
  geom_qq() +
  facet_wrap(~age_banded)
```


```{r}
ggplot(data = warblers_winglength_mass, aes(x = wing_r, y = mass, color = sex)) +
  geom_point() +
  facet_wrap(~age_banded)
#A visualization (ggplot() + geom_point()) of warbler wing length (in millimeters) in relation to bird mass (in grams), with color (aes(, color = sex)) indicating bird sex and the graphs faceted (facet_wrap(~age_banded)) by age of bird at banding.

  
```

```{r, include=FALSE}
ggplot(data = warblers_filtered, aes(x = wing_r, y = mass, color = sex)) +
  geom_point() +
  facet_wrap(~age_banded)
#Same visualization as above, just done with warblers that had all data collected.
```


### Visualization II

A visualization of warbler bill depth (in millimeters) relative to wing length (in millimeters).
```{r}
#Visualization using geom_boxplot() and geom_jitter(), with color indicating sex and plots faceted by elevation of capture

ggplot(data = warblers_filtered, aes(x = tarsus1, y = mass, color = sex)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.25) +
  facet_wrap(~elevation)


```

```{r, include = FALSE}
#Visualization using geom_boxplot() and geom_jitter(), with color indicating sex and plots faceted by elevation of capture
ggplot(data = warblers_filtered, aes(x = bill_depth, y = mass, color = sex)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.25) +
  facet_wrap(~elevation)
```


### Visualization III

A visualization of warbler bill culmen length (in millimeters) relative to bid mass (in grams).

```{r, warning=FALSE}
#Visualization using geom_violin() and geom_point(), with color indicating sex and plots faceted by elevation of bird capture
ggplot(data = warblers_filtered, aes(x = bill_culmen, y = tarsus1, color = sex)) +
  geom_violin() +
  geom_point(alpha = 0.25) +
  facet_wrap(~elevation)
  
```

## Statistical Analysis and Diagnostics

### Model Fit

GGpairs() is used to explore distributions of various variables in the warbler data to ensure that data is normally distributed

```{r, include = FALSE}
#Ggpairs() is used to examine warbler data in relation to all variables. 
warblers_filtered %>% 
  select(sex:wing_r) %>% 
  ggpairs(aes(color = sex))

#Ggpairs() is used to examine warbler data in relation to numeric variables. 
warblers_filtered %>% 
  select(sex, bill_depth:wing_r) %>% 
  ggpairs(aes(color = sex))
```

### Akaike Information Criterion (AIC)

Multiple linear regression is performed under various scenarios and then models are assessed for fit using AIC

```{r, include = FALSE}

warblerLM <- lm(mass ~ wing_r + sex, data = warblers_filtered) # Model with wing length and sex
AIC(warblerLM)


warblerLM2 <- lm(mass ~ wing_r + sex + elevation + age_banded, data = warblers_filtered) # Model with wing length, sex, elevation of capture, and age at banding.
AIC(warblerLM2)

warblerLM3 <- lm(mass ~ wing_r + sex + age_banded, data = warblers_filtered) # Model with wing length, sex, and age at banding
AIC(warblerLM3)

warblerLM4 <- lm(mass ~ wing_r + sex + elevation, data = warblers_filtered) # Model with wing length, sex, and elevation of capture
AIC(warblerLM4)

warblerLM5 <- lm(mass ~ wing_r + sex + bill_depth + bill_culmen + tarsus1, data = warblers_filtered) # Model with wing length, sex, bill depth, culmen length, and tarsus length.
AIC(warblerLM5)

warblerLM6 <- lm(mass ~ wing_r + sex + bill_depth + bill_culmen, data = warblers_filtered) # Model with wing length, sex, bill depth, and culmen length.
AIC(warblerLM6)

warblerLM7 <- lm(mass ~ wing_r + sex + bill_culmen + tarsus1, data = warblers_filtered) # Model with wing length, sex, culmen length, and tarsus length.
AIC(warblerLM7)

warblerLM8 <- lm(mass ~ wing_r + sex + tarsus1, data = warblers_filtered) # Model with wing length, sex, and tarsus length.
AIC(warblerLM8)

warblerLM9 <- lm(mass ~ wing_r + sex + bill_depth, data = warblers_filtered) # Model with wing length, sex, and bill depth.
AIC(warblerLM9)

warblerLM10 <- lm(mass ~ wing_r + sex + bill_culmen, data = warblers_filtered) # Model with wing length, sex, and culmen length.
AIC(warblerLM10)

warblerLM11 <- lm(mass ~ wing_r + sex + bill_depth + tarsus1, data = warblers_filtered)# Model with wing length, sex, bill depth, culmen length, and tarsus length.
AIC(warblerLM11)
```

### Model Assumptions and Diagnostics

Assess model fit, residuals, and assumptions using plot() function for diagnostics 

```{r}
#
AICwarblerLM7 <- AIC(warblerLM7)
AICwarblerLM7
summary(warblerLM7)
plot(warblerLM7)
vif(warblerLM7)

warblerLM7_tidy <- tidy(warblerLM7)
```

### Multiple Linear Regression

```{r, results = "asis"}
stargazer(warblerLM7, type = "html")
#
```
Multiple linear regression is performed to create a model for warbler mass in relation to four variable: wing length (in millimeters), sex (male or female), culmen length (in millimeters), and tarsus length (in millimeters).

-The model was assessed for fit in multiple ways, with ggpairs() indicating a normal distribution of all variables.

-AIC was used to determine the proper balance of model fit and variable, with the above mentioned multiple linear regression model yielding the lowest Akaike Information Criterion value (AIC = 1191.78). AIC is a relative measure and is useful in comparing models but needs to b used in conjunction with other assessments.

-When using pairs() to assess the assumptions of the model, all assumptions were met (residuals vs. fitted values/homoscedasticity, normal distribution of the residuals, and cook's distance of residuals) although there is one warbler data point that is an outlier, a male warbler with a mass of 14.5 grams. Should also be conscious about interaction between predictor variables. Variance Inflation Factor (VIF) values show this to be less likely

### Discussion of Final Model
Mass = `r round(warblerLM7_tidy$estimate[1], digits = 3)` + `r round(warblerLM7_tidy$estimate[2], digits = 3)`(wing_r)  `r round(warblerLM7_tidy$estimate[3], digits = 3)`(Male) + `r round(warblerLM7_tidy$estimate[4], digits = 3)`(bill_culmen) + `r round(warblerLM7_tidy$estimate[5], digits = 3)`(tarsus1)

- **wing_r:** The coefficient of `r round(warblerLM7_tidy$estimate[2], digits = 3)` for wing length (millimeters) indicates that on average, we expect a `r round(warblerLM7_tidy$estimate[2], digits = 3)` gram increase in warbler mass for each additional millimeter in right wing length.

- **sex male:** The coefficient of `r round(warblerLM7_tidy$estimate[3], digits = 3)` for males indicates that on average, if all other variables are equal, a male warbler's mass to be 1.034 less than a female warbler. Female is the reference sex.

- **bill_culmen:** The coefficient of `r round(warblerLM7_tidy$estimate[4], digits = 3)` for for bill culmen length (millimeters) indicates that on average, we expect a `r round(warblerLM7_tidy$estimate[4], digits = 3)` gram increase in warbler mass for each additional millimeter of bill culmen length. The culmen goes from the tip of the bill to the feathers at top of head.

- **tarsus1 Value:**The coefficient of `r round(warblerLM7_tidy$estimate[5], digits = 3)` for tarsus length (millimeters) indicates that on average, we expect a `r round(warblerLM7_tidy$estimate[5], digits = 3)` gram increase in warbler mass for each additional millimeter of tarsus (foot) length.

- **Intercept Value:** The coefficient value (`r round(warblerLM7_tidy$estimate[1], digits = 3)`) means that a female warbler with zero values for all the continuous explanatory variables would have a mass of `r round(warblerLM7_tidy$estimate[1], digits = 3)` grams, which is not a very useful statement.


## Works Cited
- Rodenhouse, N., R. Holmes, S. Sillett, and M. Webster. 2019. Black-throated Blue Warbler capture histories, Hubbard Brook Experimental Forest, 1986 - 2010 ver 4. Environmental Data Initiative. https://doi.org/10.6073/pasta/ea985022acfdb76cd7beeaf8da6c39b1
