---
name: Aaron
surname: Kreisberg
position: "Master's Student"
address: "Bren School of Environmental Science & Management"
phone: +805 679 1578
www: https://aaronkreisberg.github.io/aaron-website/
email: "aaronkreisberg@ucsb.edu"
twitter: https://twitter.com/ajkberg90
github: https://github.com/aaronkreisberg
linkedin: https://www.linkedin.com/in/aaron-kreisberg-84358a1bb/
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::latexcv:
    theme: classic
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```


# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, 
  "Bachelor of Science, Hydrological Sciences & Policy & Bachelor of Arts, History of Public Policy", "2012", "University of California, Santa Barbara",
  "Master of Environmental Science and Management ", "2022", "Bren School of Environmental Science & Management – University of California, Santa Barbara (UCSB)"
) %>% 
  detailed_entries(Degree, Year, Institution)
```

# Management Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

# Conservation Field Work Experience

```{r}
library(dplyr)
knitr::write_bib(c("vitae", "tibble"), "packages.bib")

bibliography_entries("packages.bib") %>%
  arrange(desc(author$family), issued)
```

# Skills, Certifications, & Travel
Computer Skills: RStudio, GitHub, GIS, Microsoft Office Suite (Word, Excel, PowerPoint), Google Suite
Certifications: Wilderness First Responder, Wilderness Medical Associates International (Exp. 9/23)
Travel: Working Holiday to New Zealand (11/19–3/20) and work in Queen Charlotte Sound (12/19–2/20)
